---
title: "Make a secret tunnel"
subtitle: "use `.ssh/config` file to connect to a remote server"
author: "Sarah Tanja"
date: 11/04/2024
categories: [code]
---

## How to setup a 'tunnel' into a remote server

Open a terminal ( I opened my WSL Ubuntu terminal 🐧) and make a .ssh directory in the home directory, and then open the nano text editor to write a config file in the .ssh folder:

```{r, engine='bash', eval=FALSE, echo=TRUE}
mkdir .ssh nano ~/.ssh/config
```

::: callout-tip
The .ssh directory is hidden because it starts with a `.`, and you won't see it unless you run `ls -a`
:::

In the nano text editor I copied the following:

```{r, engine='bash', eval=FALSE, echo=TRUE}
# secret tunnel     
Host <server>_tunnel         
    HostName <123.45.678.90>         
    User <username>         
    ControlMaster auto         
    ServerAliveInterval 30         
    ServerAliveCountMax 1200         
    ForwardX11 yes         
    LocalForward localhost:9000 127.0.0.1:9000
```

where:

-   **\<server\>** is the remote server name (make this short and sweet, since you will execute this as a command later on and must needs type it out)

-   **\<123.45.678.90\>** is the remote server IP address in the HostName field

-   **\<username**\> is the username setup for the remote server in the User field

I saved the config file by writing it out (Ctrl+X) and pressing Enter.

Next, I tested the tunnel by opening a new Linux Ubuntu terminal and running:

```{r, engine='bash', eval=FALSE, echo=TRUE}
ssh <server>_tunnel
```

In addition, you can test direct access to the remote server (not using the tunnel) with:

```{r, engine='bash', eval=FALSE, echo=TRUE}
ssh <username>@<IP-address>
```

The tunnel connected me to the remote server and prompted me to enter the password associated with my user on the remote server.

::: callout-tip
Once tunneled into the remote server you'll notice the \<username\>\@\<server\> in the terminal changes from your local computer to the remote server.
:::

You can go back and check on the .ssh/config file by navigating to the .ssh directory and running `cat` on the config file. `cat` is useful for small text files, it displays the contents directly in the terminal.

::: callout-caution
Careful about calling `cat` on really large files because everything is printed at once and the output will scroll past very quickly (and often this is not useful)
:::

```{r, engine='bash', eval= FALSE, echo=TRUE}
cat config
```
