---
title: "23-Aug-2024 Embryo RNA Extractions"
author: "Sarah Tanja"
date: '23/08/2024'
categories: [lab records]
draft: false
toc: true
toc-exand: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
link-external-icon: true
link-external-newwindow: true
---

# Samples

-   4L14F, 5CL14F, 6L14HF = 456L14

-   4M14F, 5M14F, 6M14F = 456M14

-   7C4F, 8C4F, 9C4F = 789C4

-   7H9F, 8H9F, 9H9F = 789H9

-   8C14F, 9C14F = 89C14F[^1]

-   10M4F, 11M4F, 12M4F = 101112M4

-   10H4F, 11H4F, 12H4F = 101112H4

-   10M9F, 11M9F, 12M9F = 101112M9

-   13H14F, 14H14F, 15H14F = 131415H14

-   13L9F, 14L9F, 15L9F = 131415L9

[^1]: No 7C14F sample

# Summary

::: callout-note
Overall success! Even in 89C14, which pooled 2 crosses instead of 3. Here I added extra DNA/RNA Shield to 89C14 to make the volume in the Zymo BashingBead for all samples uniform. AND despite forgetting to first run DNA/RNA Wash Buffer through the RNA filter prior to DNase treatment.
:::

# Extraction Notes

This uses the Zymo Quick-DNA/RNA Miniprep Plus Kit and unless otherwise specified here, follows the [M. Capitata Embryo DNA/RNA Extractions Protocol](https://sarahtanja.github.io/quarto-blog/posts/projects/coral-embryo-leachate/embryo-extractions.html). All centrifugation ocurred at 14800rpm/16,162 xg.

-   Thaw samples to room temp

-   Transfer and pool samples to one Zymo BashingBead Tube

    -   89C14 I 'topped up' with DNA/RNA Shield so that it had 1.5mL volume same as the other samples

-   Homogenize for 10 minutes

    -   10 min Zymo BashingBead Tube

        -   The 4hpf samples are hardier and take a little longer to lyse than the other embryonic phases

-   Centrifuge for 1 min to get rid of bubbles

    -   Check for complete lysis of embryos
    -   Supernatent is clear!

-   Transfer 2 550uL aliquots of cleared supernatent to one new nuclease-free tube

-   Centrifuge for 1 min to pellet any beads

-   Split cleared supernatent into two aliquots of 530uL & transfer each aliquot to a new nuclease-free tube

-   Add 530uL of DNA/RNA Lysis Buffer to each tube ( a total of 1060uL volume in each tube )

-   Vortex to mix for 5 seconds

    ---\> Proceed to purification

-   Sequentially transfer each sample in aliquots of 700uL into yellow Spin-Away Filter in a Collection Tube and centrifuge for 30s. **SAVE THE FLOW THROUGH FOR RNA PURIFICATION by tipping flow through into spare collection tube (15mL Falcon Tube). This will take 3 passes.**

-   Doubled volume of 100% (200 proof) ethanol to the RNA flow through in a falcon tube and vortexed to mix. End volume of flow through 2.1mL + ethanol 2.1mL = 4.2mL

-   Sequentially transfer RNA flow-through + ethanol in volumes =\>700uL into the green Zymo Spin IIICG Column in a Collection Tube. This should take 6 'spins'

-   Saved 1st 2 spins (1400uL) for protein tube. Once collected place in -20 freezer.

-   Performed DNase I Treatment on green RNA spin column

-   ::: callout-warning
    Forgot to do DNA/RNA Wash 400uL prior to DNase Treatment :(
    :::

    ```{r}
    DB = 75*10
    print(DB)
    Dnase = 5*10 
    print(Dnase)
    ```

::: callout-note
```         
    I didn't invert to mix... I always lose some volume 'beads' in the lid when I do this. Instead I flicked the tube to mix it and kept all the volume down in the conical end of the tube. This worked well and I had a full 80uL of treatment per spin column to work with!
```
:::

-   400uL of Prep Buffer
-   700uL of DNA/RNA Wash Buffer
-   400uL of DNA/RNA Wash Buffer, 2min
-   Centrifuge 2min dry run
-   Warmed Zymo DNase-RNase Free Water to 55C in heat block before elution, and slowly dripped 50uL of water directly over filter.

# End Products

eluted DNA volume: 50uL in `Zymo DNase-RNase Free Water`

eluted RNA volume: 50uL in warmed (55C) `Zymo DNase-RNase Free Water`

::: callout-important
I upped the elution volume from 30 to 50 for RNA because I wanted to make sure the nuclease-free water volume was enough to saturate the whole filter and get the RNA from it... The minimum volume in the kit instructions is 50uL... And I do have an RNA Clean & Concentrator kit that allows me to reduce that if needed. As long as total RNA is at least 100ng, we're good. That means that qubit value (ng/uL) times the elution volume.
:::

# Qubit QAQC

Using the Invitrogen Thermo Fischer RNA High Sensitivity Assay Kit.

Prepared working solution for 6 samples and two standards (plus some extra) in a 5mL tube:

```{r}
buffer <- 199*15
print(buffer)
reagent <- 1*15
print(reagent)
```

### RNA

standard 1: 96.18

standard 2: 1912.96

RunID: 2024-08-24_xxxxxx[^2]

[^2]: The Qubit date is ahead by 1 day (and I haven't figured out how to set it straight yet)

+-------------------+-------------+-------------+
| sample_id         | qubit_rna_1 | qubit_rna_2 |
+===================+=============+=============+
| -   456L14        | 5.20 ng/uL  | 5.22 ng/uL  |
+-------------------+-------------+-------------+
| -   456M14        | 5.00 ng/uL  | 5.00 ng/uL  |
+-------------------+-------------+-------------+
| -   789C4         | 7.52 ng/uL  | 7.60 ng/uL  |
+-------------------+-------------+-------------+
| -   789H9         | 7.70 ng/uL  | 7.76 ng/uL  |
+-------------------+-------------+-------------+
| -   89C14         | 5.06 ng/uL  | 5.02 ng/uL  |
+-------------------+-------------+-------------+
| -   101112M4      | 12.5 ng/uL  | 12.6 ng/uL  |
+-------------------+-------------+-------------+
| -   101112H4      | 8.00 ng/uL  | 8.10 ng/uL  |
+-------------------+-------------+-------------+
| -   101112M9      | 7.60 ng/uL  | 7.62 ng/uL  |
+-------------------+-------------+-------------+
| -   131415H14     | 6.12 ng/uL  | 6.22 ng/uL  |
+-------------------+-------------+-------------+
| -   131415L9      | 9.70 ng/uL  | 9.70 ng/uL  |
+-------------------+-------------+-------------+

# Storage Location

RNA samples are stored in -80C 'JPG Lab Thermo' freezer on shelf 1 (top shelf) in `coral-embryo-leachate RNA` wax freezer boxes with green label tape.

DNA samples (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate DNA` wax freezer box with yellow label tape.

Protein flow-through (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate PROTEIN` wax freezer box with blue label tape.
