---
title: "19-Aug-2024 Embryo RNA Extractions"
author: "Sarah Tanja"
date: '19/08/2024'
categories: [lab records]
draft: false
toc: true
toc-exand: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
link-external-icon: true
link-external-newwindow: true
---

# Samples

-   4H9F, 5H9F, 6H9F = 456H9

-   1H4F, 2H4F, 3H4F = 123H4

-   7L14F, 8L14F, 9L14F = 789L14

-   10C14F, 11C14F, 12C14F = 101112C14

-   13M9F, 14M9F, 15M9F = 131415M9

-   13L4F, 14L4F, 15L4F = 131415L4

# Summary

::: callout-note
Overall a success. Some samples have low quantity, but enough total RNA to be able to sequence. The highest RNA quantity was in the sample pooled for initial cleavage (4 hpf) that had taken longer to lyse... and ended up being homogenized in the ZYmo BashingBead Tube for a total of 5 minutes. The lower quantity samples were from 9 hpf and 14 hpf phase samples. These samples tend to have less starting material (often because by this time, there were fewer embryos that had 'survived' to these later stages).
:::

# Extraction Notes

kit: Zymo Quick-DNA/RNA Miniprep Plus Kit

-   Thaw samples to room temp

-   Transfer and pool samples to one Zymo BashingBead Tube

-   Homogenize for 3 minutes

    -   3 min Zymo BashingBead Tube

        -   The 4hpf samples are hardier and take a little longer to lyse than the other embryonic phases

-   Centrifuge at 16,000xg for 30 seconds to get rid of bubbles

    -   Check for complete lysis of embryos
    -   Incomplete lysis of 123H4

-   Homogenized for 2 more minutes

-   Centrifuged at 16,000xg for 1 min

    -   H4 still showing intact eggs!

-   Homogenized H4 only for another 3 minutes

-   Transfer 2 550uL aliquots of cleared supernatent to one new nuclease-free tube

-   Centrifuge at 16,000xg for 1 min to pellet any beads

-   Split cleared supernatent into two aliquots of 530uL & transfer each aliquot to a new nuclease-free tube

-   Add 530uL of DNA/RNA Lysis Buffer to each tube ( a total of 1060uL volume in each tube )

-   Vortex to mix for 5 seconds

    ---\> Proceed to purification

-   Sequentially transfer 1000uL of each sample in aliquots of 700uL into yellow Spin-Away Filter in a Collection Tube and centrifuge for 30s at 16,000xg **SAVE THE FLOW THROUGH FOR RNA PURIFICATION by tipping flow through into spare collection tube (15mL Falcon Tube)**

-   Doubled volume of 100% (200 proof) ethanol to the RNA flow through in a falcon tube and vortexed to mix. End volume of flow through + ethanol = 4mL

-   Sequentially transfer RNA flow-through + ethanol in volumes =\>700uL into the green Zymo Spin IIICG Column in a Collection Tube. This should take 6 'spins'

-   Saved 1st 2 spins (1400uL) for protein tube. Once collected place in -20 freezer.

-   Performed DNase I Treatment on green RNA spin column

    ```{r}
    DB = 75*6
    print(DB)
    Dnase = 5*6 
    print(Dnase)
    ```

::: callout-note
```         
    I didn't invert to mix... I always lose some volume 'beads' in the lid when I do this. Instead I flicked the tube to mix it and kept all the volume down in the conical end of the tube. This worked well and I had a full 80uL of treatment per spin column to work with!
```
:::

-   400uL of Prep Buffer
-   700uL of DNA/RNA Wash Buffer
-   400uL of DNA/RNA Wash Buffer, 2min
-   Centrifuge 2min dry run
-   Warmed Zymo DNase-RNase Free Water to 55C in heat block before elution, and slowly dripped 50uL of water directly over filter.

# End Products

eluted DNA volume: 50uL in `Zymo DNase-RNase Free Water`

eluted RNA volume: 50uL in warmed (55C) `Zymo DNase-RNase Free Water`

::: callout-important
I upped the elution volume from 30 to 50 for RNA because I wanted to make sure the nuclease-free water volume was enough to saturate the whole filter and get the RNA from it... The minimum volume in the kit instructions is 50uL... And I do have an RNA Clean & Concentrator kit that allows me to reduce that if needed. As long as total RNA is at least 100ng, we're good. That means that qubit value (ng/uL) times the elution volume.
:::

# Qubit QAQC

Using the Invitrogen Thermo Fischer RNA High Sensitivity Assay Kit.

Prepared working solution for 6 samples and two standards (plus some extra) in a 5mL tube:

```{r}
buffer <- 199*10
print(buffer)
reagent <- 1*10
print(reagent)
```

### RNA

standard 1: 105.99

standard 2: 1962.49

RunID: 2024-08-20_060012

+---------------+-------------+-------------+
| sample_id     | qubit_rna_1 | qubit_rna_2 |
+===============+=============+=============+
| -   456H9     | 10.6 ng/uL  | 10.5 ng/uL  |
+---------------+-------------+-------------+
| -   123H4     | 20.6 ng/uL  | 20.4 ng/uL  |
+---------------+-------------+-------------+
| -   789L14    | 4.60 ng/uL  | 4.40 ng/uL  |
+---------------+-------------+-------------+
| -   101112C14 | 4.20 ng/uL  | 4.20 ng/uL  |
+---------------+-------------+-------------+
| -   131415M9  | 4.80 ng/uL  | 4.80 ng/uL  |
+---------------+-------------+-------------+
| -   131415L4  | 14.1 ng/uL  | 14.0 ng/uL  |
+---------------+-------------+-------------+

# Storage Location

RNA samples are stored in -80C 'JPG Lab Thermo' freezer on shelf 1 in `coral-embryo-leachate RNA` wax freezer boxes with green label tape:

DNA samples (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate DNA` wax freezer box with yellow label tape:

Protein flow-through (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate PROTEIN` wax freezer box with blue label tape:
