---
title: "16-Aug-2024 Embryo RNA Extractions"
author: "Sarah Tanja"
date: '16/08/2024'
categories: [lab records]
draft: false
toc: true
toc-exand: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
link-external-icon: true
link-external-newwindow: true
---

# Samples

-   1L4F, 2L4F, 3L4F = 123L4

    -   1 min Zymo BashingBead Tube

-   7L9F, 8L9F, 9L9F = 789L9

    -   1 min Zymo BashingBead Tube

-   10C9F, 11C9F, 12C9F = 101112C9

    -   1 min Zymo BashingBead Tube

-   10H9F, 11H9F, 12H9F = 101112H9

    -   1 min Zymo BashingBead Tube

-   13M14F, 14M14F, 15M14F = 131415M14

    -   1 min Zymo BashingBead Tube

-   13C4F, 14C4F, 15C4F = 131415C4

    -   1 min Zymo BashingBead Tube

# Summary

::: callout-note
:::

# Extraction Notes

date: 16-AUG-2024

kit: Zymo Quick-DNA/RNA Miniprep Plus Kit

-   Thaw samples to room temp

-   Vortex to mix

-   Transfer to Zymo BashingBead Tube

-   Homogenize for 3 minutes

    -   The 4hpf samples are hardier and take a little longer to lyse than the other embryonic phases

-   Centrifuge at 16,000xg for 30 seconds to get rid of bubbles

    -   Check for complete lysis of embryos

-   Transfer 2 550uL aliquots of cleared supernatent to a new nuclease-free tube

-   Centrifuge at 16,000xg for 1 min to pellet any beads

-   Split cleared supernatent into two aliquots of 500uL & transfer each aliquot to a new nuclease-free tube

-   Added 500uL of DNA/RNA Lysis Buffer to each tube ( a total of 1000uL volume in each tube )

-   Vortex to mix for 5 seconds

    ---\> Proceed to purification

-   Sequentially transfer 1000uL of each sample in aliquots of 700uL into yellow Spin-Away Filter in a Collection Tube and centrifuge for 30s at 16,000xg **SAVE THE FLOW THROUGH FOR RNA PURIFICATION by tipping flow through into spare collection tube (15mL Falcon Tube)**

-   Doubled volume of 100% (200 proof) ethanol to the RNA flow through in a falcon tube and vortexed to mix. End volume of flow through + ethanol = 4mL

-   Sequentially transfer RNA flow-through + ethanol in volumes =\>700uL into the green Zymo Spin IIICG Column in a Collection Tube. This should take 6 'spins'

-   Saved 1st 2 spins (1400uL) for protein tube. Once collected place in -20 freezer.

-   Performed DNase I Treatment on green RNA spin column

    ::: callout-caution
    Forgot to run DNA/RNA Wash prior to DNase I treatment!!!
    :::

    ```{r}
    DB = 75*6
    print(DB)
    Dnase = 5*6 
    print(Dnase)
    ```

    -   I didn't invert to mix... I always lose some volume 'beads' in the lid when I do this. Instead I flicked the tube to mix it and kept all the volume down in the cnical end of the tube. This worked well and I had a full 80uL of treatment per spin column to work with.

-   Warmed Zymo DNase-RNase Free Water to 55C in heat block before elution, and slowly dripped water directly over filter.

eluted DNA volume: 30uL in `Zymo DNase-RNase Free Water`

eluted RNA volume: 50uL in warmed (55C) `Zymo DNase-RNase Free Water`

::: callout-important
I upped the elution volume from 30 to 50 for RNA because I wanted to make sure the nuclease-free water volume was enough to saturate the whole filter and get the RNA from it... The minimum volume in the kit instructions is 50uL... And I do have an RNA Clean & Concentrator kit that allows me to reduce that if needed. As long as total RNA is at least 100ng, we're good. That means that qubit value (ng/uL) times the elution volume.
:::

# Qubit

Using the Invitrogen Thermo Fischer RNA High Sensitivity Assay Kit.

Prepared working solution for 6 samples and two standards (plus some extra) in a 5mL tube:

```{r}
buffer <- 199*10
print(buffer)
reagent <- 1*10
print(reagent)
```

### RNA

standard 1: 96.43

standard 2: Error... re-run cal standards? No...

RunID: 2024-08-17_045016

| sample_id | qubit_rna_1 | qubit_rna_2 |
|-----------|-------------|-------------|
| 123L4     | 10.9 ng/uL  | 10.9 ng/uL  |
| 789L9     | 6.32 ng/uL  | 6.26 ng/uL  |
| 101112C9  | 5.56 ng/uL  | 5.46 ng/uL  |
| 101112H9  | 8.46 ng/uL  | 8.46 ng/uL  |
| 131415M14 | 7.38 ng/uL  | 7.32 ng/uL  |
| 131415C4  | 11.1 ng/uL  | 11.1 ng/uL  |

# Storage Location

RNA samples are stored in -80C 'Old Friedman' Sanyo freezer on shelf 1 in `coral-embryo-leachate RNA` wax freezer boxes with green label tape:

DNA samples (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate DNA` wax freezer box with yellow label tape:

Protein flow-through (not quantified) are stored in -20C in JPG Lab FSH 236 in shelf 2 in `coral-embryo-leachate PROTEIN` wax freezer box with blue label tape:
