---
title: "Field collection at Constellation Park, Alki, West Seattle"
author: "Sarah Tanja"
date: '03/30/2024'
categories: [anemones, lab records]
draft: false
toc: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
citation-location: margin
reference-location: margin
footnotes-hover: true
link-external-icon: true
link-external-newwindow: true
---

On Saturday, March 30th, 2024 Kip and I collected *Anthopleura elegantissima* from Constellation Park near Alki in West Seattle, Washington under WDFW scientific collection permit Tanja 24-010. At this site we collected 10 anemones from each of 4 aggregates, a total of 40 anemones. However, during this collection we didn't bother with keeping track of the aggregates. These anemones were collected for use in the preliminary respiration and photosynthesis response curves, and we were not concerned with genotype since we wouldn't be subjecting them to RNA extraction and differential gene expression analysis.

We made the collections at 15:00, at a -0.4ft low tide. Thankfully, *Anthopleura elegantissima* are found higher in the intertidal zone. We used metal scoopulas to gently pry the anemones pedal disc from the rocky substrate, same as on [18MAR2024 at Owens Beach](18MAR24-owens-beach.qmd).

::: column-margin

:::

```{r echo=TRUE, message=FALSE, warning=FALSE}
## install packages
if (!require("tidyverse", character.only = TRUE)) { install.packages("tidyverse")}
if (!require("kableExtra", character.only = TRUE)) { install.packages("kableExtra")}
if (!require("leaflet", character.only = TRUE)){ install.packages("leaflet")}
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## load libraries
library(tidyverse)
library(kableExtra)
library(leaflet)
```

## Data sheet

```{r}
anemone_collection <- read_csv('30MAR24-anemone-collections.csv',
                               col_types = cols(
                                 date = col_date(format = "%m/%d/%Y"),
                                 timePST = col_time(format = "%H:%M"),
                                 site = col_character(),
                                 aggregate = col_factor(),
                                 lat = col_double(),
                                 lon = col_double()
                               ))
```


```{r}
anemone_collection %>%
  kbl() %>%
  kable_paper("hover", full_width = F)



## Collection map

```{r}

# Prepare the text for the tooltip:
mytext <- paste(
   "Aggregate: ", anemone_collection$aggregate, "<br/>", 
   "Date: ", anemone_collection$date) %>%
  lapply(htmltools::HTML)

# Final Map
pdp <- leaflet(anemone_collection) %>% 
  addTiles()  %>% 
  setView( lng = -122.4135, lat = 47.5719, zoom = 18 ) %>% 
  addProviderTiles("Esri.WorldImagery") %>% 
  addCircleMarkers(~lon, ~lat, 
    fillColor = "orange", fillOpacity = 0.7, color="white", radius=8, stroke=FALSE,
    label = mytext,
    labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")
  )

pdp
```

## Takeaways

-   We avoided collecting anemones in the sand, and those were more broadly scattered and not as easily defined as an 'aggregate'.
-   All anemones are expected to be hosting the brown algae *Breviolium muscatinei*, since they are golden-brown in hue and were found in places that are often exposed to light
